---
title: "GC-MS metabolomics of B. grandiflora - Spectral deconvolution"
author: "Jefferson Pastuna"
date: "2025-03-25"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

### Introduction

Write an Introduction.

### Before to start

Write a package description.

### eRah package workflow

How to star.

```{r echo=TRUE, message=FALSE, warning=FALSE}

# eRah package installation
#install.packages('erah')
# eRah library call
library(erah)

```

Folder files

```{r echo=TRUE}

# Delete all file that are not in folders
unlink('Data/eRah/*')
# Data folder path
createdt('Data/eRah/')

```

New experiment.


```{r echo=TRUE}

# Loading (*.CDF) chromatograms
instrumental <- read.csv('Data/eRah_metadata/GCMS_inst_data_CDF.csv', sep = ";")
# Loading metadata of the chromatograms (phenotype file)
phenotype <- read.csv('Data/eRah_metadata/GCMS_pheno_data.csv', sep = ";")
# Merge of metadata information with chromatograms
ex <- newExp(instrumental = instrumental,
             phenotype = phenotype,
             info = 'B. grandiflora metabolomics')

```

Compound deconvolution

Parameter

```{r echo=TRUE}

# Deconvolution parameters for chromatogram analysis between 3.01 and 43.0 min
dec.par.3to43 <- setDecPar(min.peak.width = 3,
                           min.peak.height = 400,
                           noise.threshold = 40,
                           avoid.processing.mz = c(30:32),
                           analysis.time = c(3.01, 43.0))
# Deconvolution parameters for chromatogram analysis between 43.0 and 53.5 min
dec.par.43to53 <- setDecPar(min.peak.width = 3,
                           min.peak.height = 500,
                           noise.threshold = 50,
                           avoid.processing.mz = c(30:32),
                           analysis.time = c(43.0, 53.5))
# Deconvolution parameters for chromatogram analysis between 53.5 and 58.0 min
dec.par.53to58 <- setDecPar(min.peak.width = 3,
                           min.peak.height = 1000,
                           noise.threshold = 100,
                           avoid.processing.mz = c(30:32),
                           analysis.time = c(53.5, 58.0))

```

parallel processing

```{r echo=TRUE}

plan(future::multisession,
     workers = 5)

```

Deconvolution

```{r echo=TRUE, warning=FALSE}

# Chromatogram deconvolution between 3.01 and 43.0 min
ex.3to43 <- deconvolveComp(ex,
                           dec.par.3to43)
# Chromatogram deconvolution between 43.0 and 53.5 min
ex.43to53 <- deconvolveComp(ex,
                            dec.par.43to53)
# Chromatogram deconvolution between 53.5 and 58.0 min
ex.53to58 <- deconvolveComp(ex,
                            dec.par.53to58)

```

Alignment

```{r echo=TRUE}

# Alignment parameters
ex.al.par <- setAlPar(min.spectra.cor = 0.85,
                      max.time.dist = 3,
                      mz.range = 50:500)
# Alignment of 3.01-43.0 min dataset
ex.3to43 <- alignComp(ex.3to43,
                      alParameters = ex.al.par)
# Alignment of 43.0-53.5 min dataset
ex.43to53 <- alignComp(ex.43to53,
                       alParameters = ex.al.par)
# Alignment of 53.5-58.0 set
ex.53to58 <- alignComp(ex.53to58,
                       alParameters = ex.al.par)

```

Missing compound recovery


```{r echo=TRUE}

# Missing compound recovery for 3.01-43.0 min dataset
ex.3to43 <- recMissComp(ex.3to43,
                        min.samples = 2)
# Missing compound recovery for 43.0-53.5 min dataset
ex.43to53 <- recMissComp(ex.43to53,
                         min.samples = 2)
# Missing compound recovery for 53.5-58.0 min dataset
ex.53to58 <- recMissComp(ex.53to58,
                         min.samples = 2)

```

Exporting alignment feature list.

```{r echo=TRUE}

# Extracting alignment feature list 3.01-43.0 min
feat_list.3to43 <- alignList(ex.3to43,
                             by.area = FALSE)
# Exporting alignment feature list
write.csv(feat_list.3to43,
          file = "Result/eRah_3to43min/B_gradiflora_eRah_Feat_list_3to43min.csv")
# Extracting alignment feature list 43.0-53.5 min
feat_list.43to53 <- alignList(ex.43to53,
                              by.area = FALSE)
# Exporting alignment feature list
write.csv(feat_list.43to53,
          file = "Result/eRah_43to53min/B_gradiflora_eRah_Feat_list_43to53.csv")
# Extracting alignment feature list 53.5-58.0 min
feat_list.53to58 <- alignList(ex.53to58,
                              by.area = FALSE)
# Exporting alignment feature list
write.csv(feat_list.53to58,
          file = "Result/eRah_53to58min/B_gradiflora_eRah_Feat_list_53to58.csv")

```

# Identification

Metabolite identification was by comparing all the spectra found against a reference database. The eRah default database was used to metabolite identification. Metabolite identification was improve by exporting all the spectra found (.msp) in eRah to NIST MS Search 2.4 software.

```{r echo=TRUE, warning=FALSE}

# Identification of 3.01-43.0 min dataset
peak_iden.3to43 <- identifyComp(ex.3to43,
                          id.database = mslib,
                          mz.range = NULL,
                          n.putative = 1)
# Identification of 43.0-53.5 min dataset
peak_iden.43to53 <- identifyComp(ex.43to53,
                          id.database = mslib,
                          mz.range = NULL,
                          n.putative = 1)
# Identification of 53.5-58.0 min dataset
peak_iden.53to58 <- identifyComp(ex.53to58,
                          id.database = mslib,
                          mz.range = NULL,
                          n.putative = 1)

```

Exporting spectra to NIST MS Search software identification with NIST-20 library.

```{r echo=TRUE}

# 3.01-43.0 min dataset
export2MSP(peak_iden.3to43,
           store.path = "Result/eRah_3to43min",
           alg.version = 2)
# 43.0-53.5 min dataset
export2MSP(peak_iden.43to53,
           store.path = "Result/eRah_43to53min",
           alg.version = 2)
# 53.5-58.0 min dataset
export2MSP(peak_iden.53to58,
           store.path = "Result/eRah_53to58min",
           alg.version = 2)

```

